rang: Reconstructing reproducible R computational environments

### 2023-03-10 paper/paper.qmd
Fix all in-text citations for PLOS One #111

There has been many papers written on how containerization solutions such as Docker can be helpful also to foster computational reproducibility of science [e.g. @nuest:2019;@peikert:2021:RDA;@boettiger:2017:IR]. Although tutorials are available [e.g. @nuest:2019], providing a declarative description of the computational environment in the form of Dockerfile is far from the standard code sharing practice. This might be due to a lack of (DevOps) skills of most scientists to create a Dockerfile [@kim:2018:E]. But there are many tools available to automate the process [e.g. @nuest:2019]. The case in point described in this paper, `rang`, is one of them. We argue that `rang` is the only easy-to-use solution available that can pin down and restore all four components without the reliance on any commercial service <!-- such as MRAN -->.

@@

`containerit` [@nuest:2019] takes the current state of the computational environment and documents it as a Dockerfile. `containerit` makes the assumption that Component A has a weak influence on computational reproducibility and therefore defaults to Linux-based Rocker base images [@boettiger:2017:IR]. In this way, it fixes Component A. But `containerit` does not pin down the exact version of R packages. Therefore, it can pin down components A, B, C, but only a part of component D. `dockta` is another containerization solution that can potentially pin down all components due to the fact that MRAN is used. But it also suffers from the same limitations mentioned above.

@@

 @cruewell:2023:WB evaluate the computational reproducibility of 14 articles published in *Psyhocological Science*. Among these articles, the paper by @hilgard:2019:NEG has been rated as having "package dependency issues".

@@

When running `dockerize()`, one can take advantage of the `materials_dir` parameter to transfer the shared materials from @hilgard:2019:NEG into the Docker image.

@@

All R scripts ran fine inside the container and the figures generated are the same as the ones in @hilgard:2019:NEG.

@@

The study by @trisovic:2022 evaluates the reproducibility of R scripts shared on Dataverse. They found that 75\% of R scripts cannot be successfully executed. Among these failed R scripts is an R script shared by @beck:2019:EGD.

@@

Using `rang`, we are able to reconstruct a computational environment with R 2.15.0 (2012-03-30) to run all code snippets published in @jurka:2012 [^speed]. For removed CRAN packages, we strongly recommend querying the Github read-only mirror of CRAN instead (https://github.com/cran). It is because in this way, the resolved system requirements have a higher chance of being correct.

@@

The oldest version of R that `rang` can support, as of writing, is R 1.3.1. `rang` is probably the only solution available that can support the 1.x series of R (i.e. before 2004-10-04). Similar to the case of `maxent` above, a Dockerfile to assemble a Docker image with `ptproc` installed can be generated with two lines of code.

@@

Suppose we have an R script, extracted from @peng:2003:MDP, called "peng.R" like this:

@@

[^random]: It is also important to note that the random number generator (RNG) of R has been changed several times over the course of the development. In this case, we are using the same generation of RNG as @peng:2003:MDP.

@@

`rang` has native support for Bioconductor packages since version 0.2. We obtained the R script `"PaperFigure.R"` from the Github repository of `Sushi` [^sushi], which generates the figure in @phanstiel:2014:S. Similar to the above case of `ptproc`, we made a completely automated BASH script to run `"PaperFigure.R"` and get the generated figure out of the container (@fig-figure2). We made no modification to `"PaperFigure.R"`.

@@

To demonstrate this point, we took the recent paper by @oser:2022:HPE. This paper was selected because 1) the paper was published in *Political Communication*, a high impact journal that awards Open Science Badges; 2) shared data and R code are available; and most importantly, 3) the shared R code is well-written. In the repository of this paper, we based on the materials shared by @oser:2022:HPE and prepared a research compendium that should have long-term computational reproducibility. The research compendium is similar to the Executable Compendium suggested by the Turing way.

@@

With this `Makefile`, one can create the Dockerfile with `make resolve`, build the Docker image with `make build`, render the RMarkdown file inside the container with `make render`, export the built Docker image with `make export`, and rebuild the exported Docker image with `make rebuild`.

@@

In this executable compendium, only the first four elements are essential. The directory `oserdocker` (116 MB) contains cached R packages, a Dockerfile, and a verbatim copy of the directory `meta-analysis/` to be transferred into the Docker image. That can be regenerated by running `make resolve`. However, having this directory preserved insures against the situations that some R packages used in the project were no longer available or any of the information providers used by `rang` for resolving the dependency relationships were not available. (Or in the rare circumstance of `rang` is no longer available.)

>>>>

There has been many papers written on how containerization solutions such as Docker can be helpful also to foster computational reproducibility of science [e.g. @nuest:2019;@peikert:2021:RDA;@boettiger:2017:IR]. Although tutorials are available [e.g. @nuest:2019], providing a declarative description of the computational environment in the form of `Dockerfile` is far from the standard code sharing practice. This might be due to a lack of (DevOps) skills of most scientists to create a `Dockerfile` [@kim:2018:E]. But there are many tools available to automate the process [e.g. @nuest:2019]. The case in point described in this paper, `rang`, is one of them. We argue that `rang` is the only easy-to-use solution available that can pin down and restore all four components without the reliance on any commercial service <!-- such as MRAN -->.

@@

`containerit` [@nuest:2019] takes the current state of the computational environment and documents it as a `Dockerfile`. `containerit` makes the assumption that Component A has a weak influence on computational reproducibility and therefore defaults to Linux-based Rocker base images [@boettiger:2017:IR]. In this way, it fixes Component A. But `containerit` does not pin down the exact version of R packages. Therefore, it can pin down components A, B, C, but only a part of component D. `dockta` is another containerization solution that can potentially pin down all components due to the fact that MRAN is used. But it also suffers from the same limitations mentioned above.

@@

Crüwell et al. [@cruewell:2023:WB] evaluate the computational reproducibility of 14 articles published in *Psyhocological Science*. Among these articles, the paper by Hilgard et al. [@hilgard:2019:NEG] has been rated as having "package dependency issues".

@@

When running `dockerize()`, one can take advantage of the `materials_dir` parameter to transfer the shared materials from Hilgard et al. [@hilgard:2019:NEG] into the Docker image.

@@

All R scripts ran fine inside the container and the figures generated are the same as the ones in Hilgard et al. [@hilgard:2019:NEG].

@@

The study by Trisovic et al. [@trisovic:2022] evaluates the reproducibility of R scripts shared on Dataverse. They found that 75\% of R scripts cannot be successfully executed. Among these failed R scripts is an R script shared by Beck [@beck:2019:EGD].

@@

Using `rang`, we are able to reconstruct a computational environment with R 2.15.0 (2012-03-30) to run all code snippets published in Jurka [@jurka:2012] [^speed]. For removed CRAN packages, we strongly recommend querying the Github read-only mirror of CRAN instead (https://github.com/cran). It is because in this way, the resolved system requirements have a higher chance of being correct.

@@

The oldest version of R that `rang` can support, as of writing, is R 1.3.1. `rang` is probably the only solution available that can support the 1.x series of R (i.e. before 2004-10-04). Similar to the case of `maxent` above, a `Dockerfile` to assemble a Docker image with `ptproc` installed can be generated with two lines of code.

@@

Suppose we have an R script, extracted from Peng [@peng:2003:MDP], called "peng.R" like this:

@@

[^random]: It is also important to note that the random number generator (RNG) of R has been changed several times over the course of the development. In this case, we are using the same generation of RNG as Peng [@peng:2003:MDP].

@@

`rang` has native support for Bioconductor packages since version 0.2. We obtained the R script `"PaperFigure.R"` from the Github repository of `Sushi` [^sushi], which generates the figure in the original paper [@phanstiel:2014:S]. Similar to the above case of `ptproc`, we made a completely automated BASH script to run `"PaperFigure.R"` and get the generated figure out of the container (@fig-figure2). We made no modification to `"PaperFigure.R"`.

@@

To demonstrate this point, we took the recent paper by Oser et al. [@oser:2022:HPE]. This paper was selected because 1) the paper was published in *Political Communication*, a high impact journal that awards Open Science Badges; 2) shared data and R code are available; and most importantly, 3) the shared R code is well-written. In the repository of this paper, we based on the materials shared by Oser et al. [@oser:2022:HPE] and prepared a research compendium that should have long-term computational reproducibility. The research compendium is similar to the Executable Compendium suggested by the Turing way.

@@

With this `Makefile`, one can create the `Dockerfile` with `make resolve`, build the Docker image with `make build`, render the RMarkdown file inside the container with `make render`, export the built Docker image with `make export`, and rebuild the exported Docker image with `make rebuild`.

@@

In this executable compendium, only the first four elements are essential. The directory `oserdocker` (116 MB) contains cached R packages, a `Dockerfile`, and a verbatim copy of the directory `meta-analysis/` to be transferred into the Docker image. That can be regenerated by running `make resolve`. However, having this directory preserved insures against the situations that some R packages used in the project were no longer available or any of the information providers used by `rang` for resolving the dependency relationships were not available. (Or in the rare circumstance of `rang` is no longer available.)
